---
title: "Pixels to Profits: The Evolution of Global Gaming"
author: "Marcus Lewis"
date: "May 14, 2026"
format: 
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: lux
execute:
  warning: false
  message: false
---

```{r setup}
# Loading required libraries
library(tidyverse)
library(plotly)
library(gganimate)
library(hrbrthemes)
library(gifski)

# Loading and Cleaning Data
vgsales <- read_csv("/Users/marcusjabari/Downloads/vgsales.csv") %>%
  filter(!is.na(Year), Year <= 2016) %>%
  mutate(Year = as.numeric(Year))


# Reshaping data for regional analysis
vgsales_long <- vgsales %>%
  pivot_longer(cols = c(NA_Sales, EU_Sales, JP_Sales),
               names_to = "Region",
               values_to = "Sales") %>%
  mutate(Region = str_replace(Region, "_Sales", ""))

# Preview the transformed data
head(vgsales_long) %>% knitr::kable()
```

## The History of Growth
The gaming industry saw a massive explosion in the mid-2000s, driven by the "Casual Gaming" revolution led by the Nintendo Wii.


##Animated Trend (1980 - 2016)
The following animation shows the "Golden Age of Growth," highlighting the dramatic spike in global units sold.

```{r}

# Preparing data for animation
annual_trend <- vgsales %>%
  group_by(Year) %>%
  summarize(Total_Sales = sum(Global_Sales))

# Creating animated plot
p_anim <- ggplot(annual_trend, aes(x = Year, y = Total_Sales, group = 1)) + # Added group = 1
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "steelblue", size = 2) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(1980, 2016, by = 5)) +
  labs(title = "Annual Global Game Sales",
       subtitle = "Year: {frame_along}",
       y = "Millions of Units",
       x = "Year") +
  transition_reveal(Year)

# To make it show up in Quarto, call the object explicitly:
animate(p_anim, renderer = gifski_renderer())

```
Market Leaders
Who dominates the sales charts? By creating an interactive visualization, we can explore the top 10 best-selling games of all time and see the impact of console bundling.

Interactive Top 10 Best Sellers
Hover over the bars to see specific sales figures and publishers.

```{r}
top_10 <- vgsales %>%
  slice_max(Global_Sales, n = 10)

p_interact <- ggplot(top_10, aes(x = reorder(Name, Global_Sales), 
                                 y = Global_Sales, 
                                 fill = Platform,
                                 text = paste("Game:", Name, 
                                              "<br>Sales:", Global_Sales, "M",
                                              "<br>Publisher:", Publisher))) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top 10 Global Best Sellers",
       x = "",
       y = "Global Sales (Millions)")

ggplotly(p_interact, tooltip = "text")
```
## Regional Insights
One of the most interesting findings is the cultural gap between North America and Japan. While the West is driven by Action and Shooter games, Japan remains the global stronghold for Role-Playing Games.

```{r}
vgsales_long %>%
  filter(Region %in% c("NA", "JP"), 
         Genre %in% c("Action", "Role-Playing", "Shooter", "Sports")) %>%
  group_by(Region, Genre) %>%
  summarize(Total_Sales = sum(Sales)) %>%
  ggplot(aes(x = Genre, y = Total_Sales, fill = Region)) +
  geom_col(position = "dodge") +
  theme_minimal() +
  labs(title = "Regional Genre Preferences",
       subtitle = "Comparing North America vs. Japan",
       y = "Total Sales (Millions)")
```

Conclusion
The data tells a clear story: the gaming industry is not a monolith. Its success is driven by hardware innovation (the Wii effect), cultural specialization (the RPG market in Japan), and blockbuster "Mega-Hits" that define entire decades.
