---
title: "Pixels to Profits"
subtitle: "The Evolution of the Global Gaming Industry"
author: "Marcus Lewis"
date: "May 4, 2026"
format: 
  revealjs:
    theme: sky
    transition: slide
    scrollable: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## The Hook: Why Video Games?

- **The Question:** How did a niche hobby in the 1980s become a \$200B+ global industry?
- **The Data:** 16,500+ records of sales from 1980 to 2020.
- **The Goal:** Today we will uncover:
  1.  Which genres drove global growth.
  2.  How regional tastes (US vs. Japan) differ.
  3.  The era of the "Mega-Hit."

## Our Data Source

- **Dataset:** Kaggle Video Game Sales (vgsales.csv).
- **Variables:** Rank, Name, Platform, Year, Genre, Publisher, and Regional Sales (NA, EU, JP, Other).
- **Suitability:** It contains enough categorical data (Genre/Platform) and numerical data (Sales) to track trends over 40 years.

## Data Transformation

To tell a clean story, we had to clean and reshape the raw data.

\`\`\`{r} library(tidyverse)

# Load data

```{r}
library(tidyverse)

vgsales <- read_csv("/Users/marcusjabari/Downloads/vgsales.csv") %>% 
  filter(!is.na(Year), Year <= 2016) # Filtering incomplete recent years
```

# Create a Long Format for Regional Analysis

```{r}
vgsales_long <- vgsales %>% pivot_longer(cols = c(NA_Sales, EU_Sales, JP_Sales), names_to = "Region", values_to = "Sales") %>% mutate(Region = str_remove(Region, "_Sales"))
```



YAML

```         
---
title: "Pixels to Profits"
subtitle: "The Evolution of the Global Gaming Industry"
author: "Your Name"
date: "May 4, 2026"
format: 
  revealjs:
    theme: sky
    transition: slide
    scrollable: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

## The Hook: Why Video Games?
* **The Question:** How did a niche hobby in the 1980s become a $200B+ global industry?
* **The Data:** 16,500+ records of sales from 1980 to 2020.
* **The Goal:** Today we will uncover:
    1.  Which genres drove global growth.
    2.  How regional tastes (US vs. Japan) differ.
    3.  The era of the "Mega-Hit."

## Our Data Source
* **Dataset:** Kaggle Video Game Sales (vgsales.csv).
* **Variables:** Rank, Name, Platform, Year, Genre, Publisher, and Regional Sales (NA, EU, JP, Other).
* **Suitability:** It contains enough categorical data (Genre/Platform) and numerical data (Sales) to track trends over 40 years.

## Data Transformation
To tell a clean story, we had to clean and reshape the raw data.

```{r}
library(tidyverse)

# Load data
vgsales <- read_csv("vgsales.csv") %>%
  filter(!is.na(Year), Year <= 2016) # Filtering incomplete recent years

# Create a Long Format for Regional Analysis
vgsales_long <- vgsales %>%
  pivot_longer(cols = c(NA_Sales, EU_Sales, JP_Sales),
               names_to = "Region",
               values_to = "Sales") %>%
  mutate(Region = str_remove(Region, "_Sales"))
```

## The "Golden Age" of Growth

The mid-2000s saw an explosion in sales. Why?

```{r}
#| output-location: fragment
# Summarizing sales by year
annual_sales <- vgsales %>%
  group_by(Year) %>%
  summarize(Total_Global = sum(Global_Sales))

ggplot(annual_sales, aes(x = Year, y = Total_Global)) +
  geom_line(color = "firebrick", size = 1.2) +
  geom_area(fill = "firebrick", alpha = 0.3) +
  theme_minimal() +
  labs(title = "Total Video Game Sales (1980-2016)", y = "Millions of Units")
```
